Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-16 paper/paper.tex
A pass on k=3
Fix
Moving stuff around

@@

In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book}\maciek{any better source? ski rental?} allows to obtain an optimal algorithm for $k=3$.
Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of an epoch (that we define precisely later).

To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located in the same cluster.
\maciek{Define component-respecting partition and use throughout the paper instead of "have components split"}

@@

  Now we estimate the cost of $\OPT(\sigma_E)$ by accounting savings from migrations.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This might allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  As $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
  Thus, by performing a single migration for the cost $\alpha$, $\OPT$ may not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
  Finally, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
  This gives us
  \begin{equation*}
    \OPT(\sigma_E) + \OPT(M)  \geq (\TAlg(\sigma_E)-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
    \geq \TAlg(\sigma_E) / 4 - 4\cdot \ell \cdot \alpha.
  \end{equation*}
  From that, we obtain $\TAlg(\sigma_E) \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
  With that, we are ready to bound the competitive ratio.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, using above bounds,
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + \TAlg(\sigma_E)}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
  \end{equation*}

@@

  Now we consider the last, unfinished phase.
  Consider the case, where the unfinished phase is also the first one.
  Then, we cannot charge $\OPT$ due to unability to partition the components.
  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
  and we follow the analysis regarding the unsaturated requests.
  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
  This way, the competitive ratio increases at most twofold.
\end{proof}

@@

\todo{Line endings (add tildes where necessary)}
\end{appendix}

>>>>

To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located in the same cluster.

@@

\maciek{TODO: find better source for rent-or-buy, e.g. ski rental paper}

In this section we show that the classic \emph{rent-or-buy} approach\cite{borodin-book} allows to obtain an optimal algorithm for $k=3$.
Upon receiving $\alpha$ requests between a pair of nodes, we collocate them in one cluster until the end of a phase (that we define precisely later).

@@

  Now we estimate the cost of $\OPT(\sigma_E)$ by accounting savings from migrations.
  By performing a swap of nodes $(u,v)$, $\OPT$ collocates $u$ with two nodes $u', u''$, and $v$ with two nodes $v'$, $v''$.
  This might allow to serve requests between $(u,u')$, $(u,u'')$, $(v,v')$ and $(v,v'')$ for free afterwards.
  As $\sigma_E$ consists of requests between unsaturated pairs, and it accounts only for external requests, there are at most $\alpha-1$ requests between each of these pairs.
  Thus, by performing a single migration for the cost $\alpha$, $\OPT$ may not pay the remote serving costs for at most $4\cdot (\alpha - 1)$ requests from $\sigma_E$.
  Finally, for serving $\sigma_E$, $\OPT$ pays at least $W \cdot \frac{\alpha}{4\cdot (\alpha-1)}$.
  This gives us
  \begin{equation*}
    \OPT(\sigma_E) + \OPT(M)  \geq (|\sigma_E|-3\cdot\ell\cdot(\alpha - 1)) \cdot \frac{\alpha}{4\cdot (\alpha-1)}
    \geq |\sigma_E| / 4 - 4\cdot \ell \cdot \alpha.
  \end{equation*}
  From that, we obtain $|\sigma_E| \geq 4\cdot(\OPT(\sigma_E)+\OPT(M)) + 16\ell \cdot \alpha$.
  Let $E := \OPT(\sigma_E) + \OPT(M)$. Finally, we are ready to bound the competitive ratio
  \begin{equation*}
    \frac{\TAlg(\sigma)}{\OPT(\sigma)} \leq \frac{10\alpha \cdot \ell + |\sigma_E|}{(\alpha + \OPT(\sigma_E) + \OPT(M))/2} \leq \frac{52\alpha\cdot\ell + 8\cdot E}{\alpha + E} \leq 52 \ell = O(\ell).
  \end{equation*}

@@

  It remains to consider the last, unfinished phase.
  Consider the case, where the unfinished phase is also the first one.
  Then, we cannot charge $\OPT$ due to unability to partition the components.
  Instead, we use the fact that \TAlg and $\OPT$ started with the same initial configuration.
  We charge $\OPT$ $\alpha$ for the first external $\alpha$ requests or a migration,
  and we follow the analysis regarding the unsaturated requests.
  If the input finished before first $\alpha$ external requests, then \TAlg is optimal.
  Now, consider the case, where there are at least two phases, then we split the cost $\alpha$ charged in the penultimate phase into last two phases, and follow the analysis regarding the unsaturated requests.
  This way, the competitive ratio increases at most twofold.
\end{proof}

@@

\todo{Line endings (add tildes where necessary)}

\end{appendix}
